---
title: "Script for robustness analysis"
output:
  html_document:
    keep_md: yes
    toc: yes
    theme: united
    code_folding: hide
---


This script shows how different COSMOS results of replicate runs were compared to infer information about the robustness of the optimization.  

# General settings

Set some knitr parameter and the working directory.

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  include = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  eval = TRUE,
  fig.width = 8 * (1 + sqrt(5)) / 2, fig.height = 8 # golden ratio
)

knitr::opts_knit$set(root.dir = "C:/Users/burtsche/Documents/COSMOS-TPP_paper")
```

## Load packages

```{r, message=F, warning =F}
library(plyr)
library(tidyverse) # ggplot2, purrr, tibble, dplyr, tidyr, stringr, readr, forcats
library("reshape2")
library(OmnipathR)
library(ggrepel)
# options(connectionObserver = NULL)
library(org.Hs.eg.db)
library(knitr)
library(ggpubr)

mutate <- dplyr::mutate
select <- dplyr::select
group_by <- dplyr::group_by
#devtools::install_github("saezlab/CARNIVAL@v1.3")

source("src/revision_COSMOS_functions.R")
```

# Load networks

 These networks were modelled as described with varying numbers of TPP proteins according to a monotony score introduced before (large), Fstatistic as the default option (medium) and a manual curation of high confidence hits (small). These we compared to networks without TPP.

```{r}
load("results/networks/220531_mergednetworkresults_robustness.RData")
```

# Plotting info 

```{r}
values_types =  c("TPP" = "mediumseagreen",
                  "Kinase" =  "darkcyan", 
                  "TF" = "#8b0a50",
                  "other" = "gray40")

values_TPPset_paper =  c("large" = "brown",
                   "medium" =  "indianred", 
                  "small" = "lightcoral",
                   "noTPP" = "peachpuff")

values_model =  c("input node modelled" =  "darkred", 
            "input node not modelled" = "indianred",
            "PKN node" = "gray80")

values_PKN_paper =  c("PKN2" =  "maroon", 
            "PKN1" = "goldenrod1")

```


# Robustness heatmap

## Extract relevant data frames from result objects

Nodes

```{r}
nodes <- map_dfr(result_list, 'nodes', .id = "networkID")%>% 
  dplyr::select(networkID, label, measured)
```

Node attributes

```{r}
nodes <- map_dfr(network_list, 'overall_node_summary', .id = "networkID") %>% 
  separate(networkID, into = c("TPP_set", "replicate", "PKN"), sep = "_", remove = F) %>% 
  left_join(nodes, by = c("networkID", "node" = "label")) %>% 
  mutate(model = gsub("input", "input node modelled", model),
         model = gsub("missed", "input node not modelled", model),
          model = gsub("additional", "PKN node", model))
```


### Heatmap

Clustering network replicates based on node activities 

```{r}
pheatmap_input2_COSMOS <- nodes %>%
  filter(model != "missed") %>% 
  mutate(conv = ifelse(measured == "inactive", -1, 0),
         conv = ifelse(measured == "active", 1, conv)) %>% 
  distinct(networkID, node, .keep_all = T)%>%
  acast(networkID ~ node, value.var = "conv", fill = 0)

annotation_cols <- data.frame(Node = colnames(pheatmap_input2_COSMOS))%>%
  left_join(distinct(nodes, node, type), by = c("Node"= "node")) %>% 
  arrange(desc(type)) %>% 
  distinct(Node, .keep_all = T) %>%
  column_to_rownames("Node")


my_palette <- c(colorRampPalette(colors = c("#67001F"))(n = 1),
              "gray91",
              c(colorRampPalette(colors = c("darkgreen"))(n = 1)))

pheatmap_input2_COSMOS <- pheatmap_input2_COSMOS[, abs(colSums(pheatmap_input2_COSMOS)) > 1]


annotation_rows <- data.frame(id = rownames(pheatmap_input2_COSMOS)) %>% 
  separate(id, into = c("TPP_set", "replicate", "PKN"), sep = "_", remove = F) %>% 
  dplyr::select(-replicate) %>% 
  column_to_rownames("id")

annotation_colors = list(
  type = values_types,
  TPP_set = values_TPPset_paper,
  PKN = values_PKN_paper)

p_heatmap_nodes2 <- pheatmap::pheatmap(pheatmap_input2_COSMOS, 
                   show_colnames = F, show_rownames = F,
                   annotation_col = annotation_cols,
                   annotation_colors = annotation_colors,
                   treeheight_col = 0, 
                   annotation_row = annotation_rows, 
                   color = my_palette,
                   legend_breaks = c(-0.8, 0, 0.8), 
                   legend_labels = c("inactive", "missed", "active"), 
                   fontsize = 10)

```




